---
title: "Examine how various aspects of car engines affect performance."
author: "Uday"
date: "2022-11-19"
output:
  pdf_document: default
  html_document: default
  word_document: default
---
##Placing packages in place and using them in the code:
```{r installing below packages which is used for Ploting}
install.packages("tinytex", repos = "http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")
install.packages("tidyr", repos = "http://cran.us.r-project.org")
install.packages("magrittr", repos = "http://cran.us.r-project.org")
install.packages("knitr", repos = "http://cran.us.r-project.org")
install.packages("glmnet", repos = "http://cran.us.r-project.org")
install.packages("leaps", repos = "http://cran.us.r-project.org")
install.packages("gvlma", repos = "http://cran.us.r-project.org")
install.packages("psych", repos = "http://cran.us.r-project.org")
install.packages("latticeExtra", repos = "http://cran.us.r-project.org")
install.packages("caret", repos = "http://cran.us.r-project.org")
library(knitr)
library(magrittr)
library(leaps)
library(gvlma)
library(glmnet)
library(psych)
library(caret)
library(latticeExtra)
library(dplyr)
library(tidyr)
```
##read from main input auto-mpg.csv file
```{r read from main input auto-mpg.csv file}
masterdf =  read.csv("C:/Users/Public/FinalProject/auto-mpg.csv")
str(masterdf)
```
##The dataframe has been modified.
```{r The dataframe has been modified.}
names(masterdf) = c("mpg","cylinder","displacement","horsepower","weight","acceleration","model_year","origin","car_name")
masterdf$horsepower[masterdf$horsepower=="?"] = NA
masterdf$horsepower = as.numeric(masterdf$horsepower)
masterdf$cylinder = as.numeric(masterdf$cylinder)
str(masterdf)
```
##separating the underlying data from the data that is transformed:
```{r separating the underlying data}
master1df = select(masterdf,mpg,displacement,horsepower,weight,acceleration)
master1df = na.omit(master1df)
kable(summary(master1df),row.names = FALSE)
```
##In the first instance, there were 300 split records, while in the second instance, there were 98 last records:
```{r first instance 300 and second instance 98}
first.instance = master1df[1:300,]
second.instance = na.omit(master1df[301:398,])
```
##Displacement model for first instance:
```{r displacement model}
firstinstance.dis = lm(mpg~displacement, data=first.instance)
summary(firstinstance.dis)
ffirstinstance.dis = summary(firstinstance.dis)
ffirstinstance.dis$r.squared                      
ffirstinstance.dis$adj.r.squared                  
coef(ffirstinstance.dis)                          
coef(firstinstance.dis)                          
```
##Horsepower model for first instance:
```{r horsepower model}
firstinstance.hrp = lm(mpg~horsepower, data=first.instance)
summary(firstinstance.hrp)
ffirstinstance.hrp = summary(firstinstance.hrp)   
ffirstinstance.hrp$r.squared                      
ffirstinstance.hrp$adj.r.squared                 
coef(ffirstinstance.hrp)				 
coef(firstinstance.hrp)				 
```
##Acceleration model for first instance:
```{r acceleration model}
firstinstance.acc = lm(mpg~acceleration, data=first.instance)
summary(firstinstance.acc) 
ffirstinstance.acc = summary(firstinstance.acc)
ffirstinstance.acc$r.squared 			
ffirstinstance.acc$adj.r.squared 		
coef(ffirstinstance.acc)				
coef(firstinstance.acc)				
```
##Weight model for first instance:
```{r Weight model }
firstinstance.wght = lm(mpg~weight, data=first.instance)
summary(firstinstance.wght) 
ffirstinstance.wght = summary(firstinstance.wght)
ffirstinstance.wght$r.squared			
ffirstinstance.wght$adj.r.squared 		
coef(ffirstinstance.wght)			
coef(firstinstance.wght)		
```
##Displacement model residual for first instance:
```{r Displacement model residual for first instance}
firstins.dist = lm(mpg~displacement, data=first.instance)
firstdfn_resdl = firstins.dist$residuals
plot(first.instance$mpg~firstdfn_resdl ,lwd=3, col="pink",main="Displacement Model residual for first instance",ylab = "mpg")
```
##Finding linear model residual of displacement for first instance:
```{r linear model residual of displacement for first instance}
qqnorm(resid(firstinstance.dis))
qqline(resid(firstinstance.dis))
plot(density(resid(firstinstance.dis)))
```
##Finding Histogram of model displacement for first instance:
```{r Histogram of model displacement for first instance}
hist(firstdfn_resdl ,prob=T,breaks=20,main="displacement model histogram",xlab="Residuals")
lines(density(firstdfn_resdl ),col="blue",lwd=3)
```

##Model of Horsepower Residual for first instance:
```{r Model of Horsepower Residual for first instance}
ffcsa_resi = firstinstance.hrp$residuals
plot(first.instance$mpg~ffcsa_resi ,lwd=3, col="pink",main="Horsepower Residual for first instance",ylab = "mpg")
```
##linear horsepower models for first instance:
```{r linear horsepower models for first instance}
qqnorm(resid(firstinstance.hrp))
qqline(resid(firstinstance.hrp))
plot(density(resid(firstinstance.hrp)))
```
##horsepower models histogram for first instance:
```{r horsepower models histogram for first instance}
hist(ffcsa_resi ,prob=T,breaks=20,main="Horsepower Histogram for first instance",xlab="Residuals")
lines(density(ffcsa_resi ),col="blue",lwd=3)
```

##acceleration residual model for first instance:
```{r acceleration residual model for first instance}
fstins_res = firstinstance.acc$residuals
plot(first.instance$mpg~fstins_res ,lwd=3, col="pink",main="Acceleration Residual for first instance",ylab = "mpg")
```
##linear acceleration residual model for first instance:
```{r linear acceleration residual model for first instance}
qqnorm(resid(firstinstance.acc))
qqline(resid(firstinstance.acc))
plot(density(resid(firstinstance.acc)))
```
##model of the acceleration histogram for first instance:
```{r model of the acceleration histogram for first instance}
hist(fstins_res ,prob=T,breaks=20,main="Acceleration Histogram model for first instance",xlab="Residuals")
lines(density(fstins_res ),col="blue",lwd=3)
```

##Weight residual model for first instance:
```{r weight residual model for first instance}
fisins_residual = firstinstance.wght$residuals
plot(first.instance$mpg~fisins_residual ,lwd=3, col="pink",main="weight residual model for first instance",ylab = "mpg")
```
##linear Weight Models for first instance:
```{r linear Weight Models for first instance}
qqnorm(resid(firstinstance.wght))
qqline(resid(firstinstance.wght))
plot(density(resid(firstinstance.wght)))
```
##histogram model weight for first instance:
```{r histogram model weight for first instance}
hist(fisins_residual ,prob=T,breaks=20,main="Histogram model weight for first instance",xlab="Residuals")
lines(density(fisins_residual ),col="blue",lwd=3)
```