---
title: "U.S. Food Imports"
author: "uday"
date: "2022-11-19"
output: html_document
---

```{r installing below packages which is used in the code}
install.packages("tidyr", repos = "http://cran.us.r-project.org")
install.packages("factoextra", repos="http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")
install.packages("knitr", repos = "http://cran.us.r-project.org")
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
library(factoextra)
library(tidyr)
library(dplyr)
library(lubridate)
library(stringr) 
library(knitr)
library(ggplot2)
```

## Project Introduction

American consumers need a wide selection of high-quality, convenient foods. The American food basket represents an increasing share of tropical crops, spices, and imported gourmet goods as Americans have become wealthier and more ethnically diverse. Seasonal and climatic factors drive U.S. imports of popular types of fruits and vegetables and tropical products, such as cocoa and coffee. Additionally, a significant portion of U.S. imports can be attributed to intra-industry trade, in which American agricultural-processing companies outsource some of their operations overseas and import goods that have undergone varying degrees of processing from their foreign-market subsidiaries.

## Data Source 

<https://catalog.data.gov/dataset/u-s-food-imports>

## Project Objective
By analyzing the aforementioned data, we will be able to determine the factual data, such as the amount of growth and income that occurred each year for each food category.

## Project Outcomes
1)To find statistics of the data.
2)Finding linear regression.
3)polynomial regression discovery.
4)Plotting the data.
5)looking for clustering in data.

```{r Actual Code}

#Reading the data into a dataframe
coredf =  read.csv("C:/Users/Public/Food_data.csv")

#The dataframe has been modified
coredf = coredf %>% rename("id" = "X","Item_Type" = "X.1","Amount_Type" = "X.2","2021" = "X2021", "2020" = "X2020", "2019" = "X2019","2018" = "X2018","2017" = "X2017","2016" = "X2016","2015" = "X2015","2014" = "X2014","2013" = "X2013","2012" = "X2012","2011" = "X2011","2010" = "X2010","2009" = "X2009","2008" = "X2008","2007" = "X2007","2006" = "X2006","2005" = "X2005","2004" ="X2004","2003" = "X2003","2002" = "X2002","2001" = "X2001","2000" = "X2000","1999" = "X1999", "Food_categories" = "Year")

#Dataframe structure before transformation 
str(coredf)

#Transforming the dataframe with structure

coredf1 = gather(coredf, key = "Year", value = "Amount", "2021", "2020", "2019","2018","2017","2016","2015","2014","2013","2012","2011","2010","2009","2008","2007","2006","2005","2004","2003","2002","2001","2000","1999")


#Renaming the columns with the the actual values
coredf1$Food_categories[coredf1$Food_categories == 'Total foods 1/'] = 'Total foods'
coredf1$Food_categories[coredf1$Food_categories == 'Subtotal foods 4/'] = 'Subtotal foods'

#Casting the values with desired types
coredf1$Year = as.numeric(coredf1$Year)
coredf1$Amount = gsub(",", "", coredf1$Amount)  
coredf1$Amount = as.double(coredf1$Amount)

#Dataframe structure after transformation 
str(coredf1)

#Redefining and removal of the NA values from the dataframe
master1df = select(coredf1,id,Year,Item_Type,Amount,Food_categories)
master1df = na.omit(master1df)

#Details of Food categories and its item types:
distinct(master1df, Food_categories,Item_Type)

#Partitioning the master dataframe based on Food categories
Total_foods = subset(master1df, Food_categories=='Total foods')
#Total Food dataframe statistics
kable(summary(select(Total_foods,id,Year,Amount)),row.names = FALSE,caption = "Total foods")

Foods = subset(master1df, Food_categories== "Foods")
#Foods dataframe statistics
kable(summary(select(Foods,id,Year,Amount)),row.names = FALSE,caption = "Foods")

Subtotal_foods = subset(master1df, Food_categories== "Subtotal foods")
#Subtotal_foods dataframe statistics
kable(summary(select(Subtotal_foods,id,Year,Amount)),row.names = FALSE,caption = "Subtotal foods")

#Finding linear regression of each Food categories

lmTotal_foods = lm(Amount~Year, data = Total_foods)
#linear regression of Total foods category
summary(lmTotal_foods)

lmall_foods = lm(Amount~Year, data = Foods)
#linear regression of Food category
summary(lmall_foods)

lmSubtotal_foods = lm(Amount~Year, data = Subtotal_foods)
#linear regression of Subtotal foods category
summary(lmSubtotal_foods)

#Polynomial regression discovery for each Food categories

#Logic of polynomial regression of Total foods category
lm(Amount ~ poly(Year, 2, raw = TRUE),data = Total_foods) 
#Logic of polynomial regression of Food category
lm(Amount ~ poly(Year, 2, raw = TRUE),data = Foods)
#Logic of polynomial regression of Subtotal foods category
lm(Amount ~ poly(Year, 2, raw = TRUE),data = Subtotal_foods)

#Plotting the data for each Food categories

#Logic for Plotting Year vs Amount of Total foods category
ggplot(Total_foods,aes(x=Year,y=Amount)) +geom_line(size=1.5) +
    labs(x = "Year", y = "Amount",caption="US$ value of Total foods imports") +
    ggtitle(paste0("Food Imports in Million $"))+
    theme_minimal()+
    theme(legend.position = "right", 
          plot.caption = element_text(hjust = 0.5),
          plot.title = element_text(hjust = 0.5,size=16),
          text = element_text(size=14))+
    scale_fill_brewer(palette="Set3") 

#Logic for Plotting Year vs Amount of Foods category
ggplot(Foods,aes(x=Year,y=Amount)) +geom_line(size=1.5) +
    labs(x = "Year", y = "Amount",caption="US$ value of Foods(Live meat animals,Meats,Fish,shellfish,Dairy,Vegetables,Fruits,Nuts,Coffee,tea,spices,Grains,Vegetableoils,Sugarcandy,Cocoa,chocolate,Other edible)") +
    ggtitle(paste0("Total Food in Million $"))+
    theme_minimal()+
    theme(legend.position = "right", 
          plot.caption = element_text(hjust = 0),
          plot.title = element_text(hjust = 0.5,size=16),
          text = element_text(size=14))+
    scale_fill_brewer(palette="Set3") 

#Logic for Plotting Year vs Amount of Subtotal foods category
ggplot(Subtotal_foods,aes(x=Year,y=Amount)) +geom_point(size=1.5) +
    labs(x = "Year", y = "Amount",caption="US$ value of Subtotal foods") +
    ggtitle(paste0("Subtotal(Animals,Plants,Beverages) foods in Million $"))+
    theme_minimal()+
    theme(legend.position = "right", 
          plot.caption = element_text(hjust = 0.5),
          plot.title = element_text(hjust = 0.5,size=16),
          text = element_text(size=14))+
    scale_fill_brewer(palette="Set3") 


#Looking for clustering in data based on for each Food categories

#Total_foods clustering
kdf = select(Total_foods,Year,Amount)
km = kmeans(kdf, centers = 4)
fviz_cluster(km, data = kdf)

#Foods clustering
kdf1 = select(Foods,Year,Amount)
km1 = kmeans(kdf1, centers = 4)
fviz_cluster(km1, data = kdf1)

#Subtotal foods clustering
kdf2 = select(Subtotal_foods,Year,Amount)
km2 = kmeans(kdf2, centers = 4)
fviz_cluster(km2, data = kdf2)

```
